Parallel Algorithms for the Probabilistic Roadmap Method for Robotic Motion Planning with GPU Acceleration

### 2016-08-04 paper/paper.tex
Abstract updates

\institute{Manuel Demmeler, Manfred Liebmann\\ 
%\at Technische Universität München\\
%Center for Mathematical Sciences, M17 \\
%Boltzmannstraße 3,
%85748 Garching by Munich \\ 
%\email{manuel.demmeler@yahoo.de}
%\and
%Manfred Liebmann
\at Technische Universität München\\
%Chair of Optimal Control,
Center for Mathematical Sciences, M17 \\
Boltzmannstraße 3,
85748 Garching by Munich \\ 
\email{manfred.liebmann@tum.de, manuel.demmeler@yahoo.de}
}

@@

\begin{abstract}
%
In trajectory optimization with obstacles for robot motion planning, it is often necessary to provide feasible initial trajectories. 
A common algorithm is the probabilistic roadmap method (PRM), which gives efficient approximative solutions and scales well with high dimensions of the underlying spaces.
%
%As the dimensions of the underlying spaces are rather high, this can be expensive. 
%A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. 
%An efficient approximative solution for this problem is the probabilistic roadmap method (PRM).
%In this paper, a parallelized version of the PRM method has been developed in general.
%A robotics application then was implemented as a special case of this with GPU support.
%
In this paper, the PRM method has been parallelized.
A first level of parallelization is achieved by GPU support.
A CUDA kernel has been implemented that runs robot kinematics calculations and collision detection for different robot configurations in parallel.
For a further level of parallelization, different approaches for the usage of multiple GPUs have been developed and benchmarked.
%
\keywords{probabilistic roadmap method \and robot motion planning \and parallel algorithms \and GPU acceleration}
\end{abstract}

@@

\begin{thebibliography}{}%{999}
	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, \url{http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html}
	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press 2005
	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press 2006
	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München 2014
	\bibitem{robodynbuch} F. Pfeiffer, E. Reithmeier, Roboterdynamik, B.G. Teubner Verlag 1987
	\bibitem{chung} K. Chung, Wenping Wang, Quick Collision Detection of Polytopes in Virtual Environments, Proceedings of the ACM Symposium on Virtual Reality Software and Technology (VRST 96), Mark Green (ed.) 1996
	\bibitem{ericson} C. Ericson, Real-Time Collision Detection, MK 2005
	\bibitem{cuda} Nvidia Corporation, Cuda Programming Guide, 2015, \url{http://docs.nvidia.com/cuda/pdf/CUDA_C_Programming_Guide.pdf}
	\bibitem{cuda2} D. B. Kirk, W. W. Hwu, Programming Massively Parallel Processors, MK 2010
	\bibitem{c++} Pitt-Francis, Guide to Scientific Computing in C++, Springer 2012
	\bibitem{mpi} W. Gropp, Using MPI: Portable Parallel Programming with the Message-Passing Interface, MIT Press 2014
	\bibitem{githubcode} M. Demmeler, Github Repository GPUPRM, \url{https://github.com/demmeler/GPUPRM}
\end{thebibliography}

>>>>

\institute{Manuel Demmeler, Manfred Liebmann\\ 
%\at Technische Universitt Mnchen\\
%Center for Mathematical Sciences, M17 \\
%Boltzmannstra§e 3,
%85748 Garching by Munich \\ 
%\email{manuel.demmeler@yahoo.de}
%\and
%Manfred Liebmann
\at Technische Universit\"at M\"unchen\\
%Chair of Optimal Control,
Center for Mathematical Sciences, M17 \\
Boltzmannstra{\ss}e 3,
85748 Garching by Munich \\ 
\email{manuel.demmeler@tum.de, manfred.liebmann@tum.de}
}

@@

\begin{abstract}
%
In trajectory optimization with obstacles for robot motion planning it is often necessary to provide feasible initial trajectories. 
A common algorithm to solve this problem is the probabilistic roadmap method (PRM), which gives efficient approximative solutions and scales well with the high dimensional spaces underlying the problem.
%
%As the dimensions of the underlying spaces are rather high, this can be expensive. 
%A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. 
%An efficient approximative solution for this problem is the probabilistic roadmap method (PRM).
%In this paper, a parallelized version of the PRM method has been developed in general.
%A robotics application then was implemented as a special case of this with GPU support.
%
In this paper, the parallelization of the PRM method specifically for graphics processing units (GPUs) is investigated.
%A first level of parallelization is achieved by GPU support.
CUDA kernels have been implemented to run the robot kinematics calculations and collision detection for different robot configurations in parallel.
Three different parallelization approaches to improve the performance for multi GPU configurations are analyzed and the resulting benchmarks are presented.
%For a further level of parallelization, different approaches for the usage of multiple GPUs have been developed and benchmarked.
%
\keywords{probabilistic roadmap method \and robot motion planning \and parallel algorithms \and GPU acceleration}
\end{abstract}

@@

\begin{thebibliography}{}%{999}
	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, \url{http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html}
	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press 2005
	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press 2006
	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl f\"ur Angewandte Mechanik, TU M\"unchen 2014
	\bibitem{robodynbuch} F. Pfeiffer, E. Reithmeier, Roboterdynamik, B.G. Teubner Verlag 1987
	\bibitem{chung} K. Chung, Wenping Wang, Quick Collision Detection of Polytopes in Virtual Environments, Proceedings of the ACM Symposium on Virtual Reality Software and Technology (VRST 96), Mark Green (ed.) 1996
	\bibitem{ericson} C. Ericson, Real-Time Collision Detection, MK 2005
	\bibitem{cuda} Nvidia Corporation, Cuda Programming Guide, 2015, \url{http://docs.nvidia.com/cuda/pdf/CUDA_C_Programming_Guide.pdf}
	\bibitem{cuda2} D. B. Kirk, W. W. Hwu, Programming Massively Parallel Processors, MK 2010
	\bibitem{c++} Pitt-Francis, Guide to Scientific Computing in C++, Springer 2012
	\bibitem{mpi} W. Gropp, Using MPI: Portable Parallel Programming with the Message-Passing Interface, MIT Press 2014
	\bibitem{githubcode} M. Demmeler, Github Repository GPUPRM, \url{https://github.com/demmeler/GPUPRM}
\end{thebibliography}
